Interactive Curvature Tensor Visualization on Digital Surfaces\thanks{This work has been mainly funded by XXXXXXX research grants.}

### 2015-10-20 latex/article.tex
intro

Volumetric objects are being more and more popular in many
applications ranging from object modeling and rendering in Computer
Graphics to geometry processing in Medical Imaging or Material
Sciences. When considering large volumetric data, interactive
visualization of such images (or isovalue surfaces) is a complex
problem. Such issues become even more difficult when dynamic
volumetric datasets are considered. Beside such isosurface
visualization, we are also interested in performing geometry
processing on the digital object and to explore different parameter
settings of the geometry processing tool.  In this work, we focus on
interactive visualization of volumetric data isosurfaces combined with
a curvature tensor (mean/Gaussian curvature, principal curvatures
directions\ldots) estimation at each point of the digital
surface. Curvature maps on a given isosurface are key information when
exploring 3D structures in Material Sciences \todo{REF}. Furthermore,
since most curvature estimators require a parameter fixing the scale
to which the computation is performed. For short, such parameter
(integration radius, convolution kernel size, \ldots) specifies noise
level on the object surface. As a consequence, exploring interactively
different values of such parameter is mandatory.

\noindent\textbf{Contributions}\quad In this work, we propose a framework
to perform interactive visualization of complex 3D digital structures
combined with a dynamic curvature tensor estimation. More precisely,
we define a fully data parallel process on GPU hardware to both
efficiently extract adaptive isosurface and compute per vertex
curvature tensor using Integral Invariants estimators. Hence, this system
allows us to visualize curvature tensor in realtime on large dynamic
objects.

\noindent\textbf{Related works}\quad Extracting and Visualizing
isosurface on volumetric data has been widelyo investigated since the
seminal Marching-Cubes approach by Lorensen and
Cline~\cite{lorensen1987marching}. This approach being data parallel,
GPU implementation of this method is very efficient
\cite{tatarchuk2007real}. However, such technique generates a lot of
triangles which is not well suited to large digital data. Hence,
adaptive approaches have been proposed in order to optimize the
triangular mesh resolution to the object geometry or depending on the
viewpoint. In this topic, many solutions have been developed in
Computer
Graphics~\cite{shu1995adaptive,schaefer2004dual,lengyel2010voxel,DBLP:journals/cgf/LewinerMPPL10,DBLP:journals/cvgip/LobelloDD14}.
The method developed by Lengyel \textit{et al.}
\cite{lengyel2010voxel} is the best suited to our needs.  However, it
was developed to visualize in real-time static or near static data and
present a complex CPU implementation.  In this paper we propose a GPU
pipeline inspired from this method.  It leaves the CPU free for other
tasks and avoids synchronisation with the GPU.  Our high framerate
also allows this method to run in real-time on fully dynamic data.
The triangulation is controlled by a GPU LoD criterion that ensures
that triangles project onto more than one pixel, thus removing
geometrical aliasing.

Curvature estimation on digital surface

>>>>

Volumetric objects are being more and more popular in many
applications ranging from object modeling and rendering in Computer
Graphics to geometry processing in Medical Imaging or Material
Sciences. When considering large volumetric data, interactive
visualization of such images (or isovalue surfaces) is a complex
problem. Such issues become even more difficult when dynamic
volumetric datasets are considered. Beside such isosurface
visualization, we are also interested in performing geometry
processing on the digital object and to explore different parameter
settings of the geometry processing tool.  In this work, we focus on
interactive visualization of volumetric data isosurfaces combined with
a curvature tensor (mean/Gaussian curvature, principal curvatures
directions\ldots) estimation at each point of the digital
surface. Curvature maps on a given isosurface are key information when
exploring 3D structures in Material Sciences \todo{REF}. Furthermore,
since most curvature estimators require a parameter fixing the scale
to which the computation is performed. For short, such parameter
(integration radius, convolution kernel size, \ldots) specifies noise
level on the object surface. As a consequence, exploring interactively
different values of such parameter is mandatory.
\vspace{0.2cm}

\noindent\textbf{Contributions}\quad In this work, we propose a
framework to perform interactive visualization of complex 3D digital
structures combined with a dynamic curvature tensor estimation. More
precisely, we define a fully data parallel process on GPU hardware to
both efficiently extract adaptive isosurface and compute per vertex
curvature tensor using Integral Invariants estimators. Hence, this
system allows us to visualize curvature tensor in realtime on large
dynamic objects. Our approach combines a GPU implementation of
pointerless octrees to represent the data, an adaptive
viewpoint-dependent mesh extraction, and a GPU implementation of
integral invariant curvature estimations.

\vspace{0.2cm}

\sloppy\noindent\textbf{Related works}\quad Extracting and Visualizing
isosurface on volumetric data has been widelyo investigated since the
seminal Marching-Cubes approach by Lorensen and
Cline~\cite{lorensen1987marching}. This approach being data parallel,
GPU implementation of this method is very efficient
\cite{tatarchuk2007real}. However, such technique generates a lot of
triangles which is not well suited to large digital data. Hence,
adaptive approaches have been proposed in order to optimize the
triangular mesh resolution to the object geometry or depending on the
viewpoint. In this topic, many solutions have been developed in
Computer
Graphics~\cite{shu1995adaptive,schaefer2004dual,lengyel2010voxel,DBLP:journals/cgf/LewinerMPPL10,DBLP:journals/cvgip/LobelloDD14}.
The method developed by Lengyel \textit{et al.}
\cite{lengyel2010voxel} is the best suited to our needs.  However, it
was developed to visualize in real-time static or near static data and
present a complex CPU implementation.  In this paper we propose a GPU
pipeline inspired from this method.  It leaves the CPU free for other
tasks and avoids synchronisation with the GPU.  Our high framerate
also allows this method to run in real-time on fully dynamic data.
The triangulation is controlled by a GPU LoD criterion that ensures
that triangles project onto more than one pixel, thus removing
geometrical aliasing.

Curvature estimation on discrete or digital surface has also been
widely investigated. In \cite{CVIU2014}, authors propose digital
versions of Integral Invariant estimators
\cite{Pottmann2007,Pottmann2009} in order to estimate the complete
curvature tensor (mean/Gaussian curvatures, principal curvatures,
principal curvature directions, normal vector field) on digital
surfaces. Such approaches are based on an integration principle using
a spherical kernel of a given radius. Additionally, authors have
demonstrated that these estimators have multigrid convergence
properties. In this article, we propose an efficient GPU
implementation of such estimators in order to visualize, in real-time,
such curvature maps and allowing us to interactively explore different
values of the spherical kernel radius.
